Learning and Reputation under Monopoly and Duopoly: \\A Competing Bandits Approach

### 2018-08-27 paper/competing_bandits.tex
tweaks

To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution. There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the\textit{relative reputation proportion}.

\begin{figure}
\caption{Reputation Distribution}
\includegraphics[scale=0.35]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
\end{figure}

@@

\begin{figure}
\caption{Reputation Difference Distribution}
\includegraphics[scale=0.35]{figures/ts_dg_rep_diff_nih}
\label{ts_dg_rep_diff_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the difference in reputation between $TS$ and $DG$ across $t$}}
\end{figure}

@@

\begin{definition}
\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative. Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm. We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
\end{definition}

\begin{figure}
\caption{Relative Reputation Plots}
\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$. The shaded area display 95\% confidence intervals.}}
\label{relative_rep_plots}

Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option. Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.

>>>>

To better capture the comparison between two algorithms we explicitly plot the distribution of the reputation difference between $TS$ and $DG$. This figure seems to confirm the intuition noted previously since the reputation distribution has its largest mass around the point just below 0 but it is skewed to the right even as $t$ gets large. Given the skewedness of the distribution, the mean is not a representative statistic of the entire reputation difference distribution. There are alternative statistics that we could consider such as the median or numerically calculating $\Pr(reputation(TS) - reputation(DG)) > 0$ using the estimated density, but instead we define a simple and natural statistic to help us reason about what happens under competition, the \textit{relative reputation proportion}.

\begin{figure}[H]
\caption{Reputation Distribution}
\includegraphics[scale=0.35]{figures/rep_distribution_nih}
\label{rep_dist_nih}
\caption*{\tiny{The plots contain a kernel density estimate of the reputation distribution at $t = 500$}}
\end{figure}

@@

\begin{figure}[H]
\caption{Relative Reputation Plots}
\includegraphics[scale=0.35]{figures/relative_uniform_annotated_plot}
\includegraphics[scale=0.35]{figures/ts_dg_nih_10_prelim}
\caption*{\tiny{The plots contain the average reputation over $1000$ runs for a memory size of $100$ where, for a given $t$, we record the reputation of both of the algorithms on a given instance and then calculate the proportion of runs where $TS \geq DG$. The shaded area display 95\% confidence intervals.}}
\label{relative_rep_plots}

@@

\begin{definition}
\textit{Relative reputation cost and benefit} - the relative reputation loss an algorithm incurs from purposeful exploration compared to the greedy alternative. Here we treat it as an empirical definition based on the relative reputation proportion where the ``cost" regime is when the proportion $< 0.5$ and the ``benefit" regime is when the proportion is $> 0.5$ for the better algorithm. We call the priors where there is an early costly period followed by a later benefit period relative reputation costly priors.
\end{definition}
Is exploration always costly? Figure \ref{relative_rep_plots} also shows that for the Needle In Haystack prior, $TS$ always does relatively better than $DG$. There are two contributing factors to this. First, $TS$ identifies the best arm faster in the Needle In Haystack prior than the Uniform prior so that there is a shorter time horizon where $TS$ needs to engage in purposeful exploration. Second, in the Needle In Haystack prior there are no ``bad" arms as there may be in the Uniform prior since by construction all the arms except one in Needle In Haystack are the same. Thus, when $TS$ pulls a sub-optimal arm relative to its current information, the expected reward is the same as the greedy option that has not identified the best arm. However, with the Uniform prior, it is possible that the sub-optimal arm that is pulled has substantially lower expected reward relative to the greedy option. Thus, only the Uniform and Heavy Tail priors are relative reputation costly priors.
